Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-07 paper/sec-mcsamples.tex
more work on the paper
More work on the paper

We consider the most relevant background
processes that contribute to the identification of
 $hh\to 4b$ candidate events.
%
This includes  QCD $4b$ multi-jet production, as well as
QCD $2b2j$ and $4j$ multi-jet samples.
%
The latter can lead to the event being identified
as signal in the case of light or charm
jets being mistagged as $b$-jets.
%
While the light jet mistag probability is small, we find that
in general the $2b2j$ and $4j$ backgrounds cannot be neglected because
of their large cross-sections and enhancement from combinatorics, that
increase their contribution as compared to a naive estimate.
%
In addition, we find that there is an important contribution from the radiation
of $b$-quarks off light partons during the parton shower, which enhances the contribution
of parton-level $2b2j$ and $4j$ events being tagged as signal events.
%

In addition to the QCD multijet processes,
we also generate $t\bar{t}$ samples
in the fully hadronic final state, which lead to a
$2b4j$ signature that can
contribute due to light jet mistags.
%
This background processes leads a similar topology that
the corresponding QCD sample.
%
Leptonic decays of top quarks can be removed by requiring
a lepton veto, and thus are not included here.
%
Single Higgs production processes, such as $Z(\to b\bar{b})h(\to b\bar{b})$
and $t\bar{t}h(\to b\bar{b})$, and electroweak backgrounds, such $Z(\to b\bar{b})b\bar{b}$,
are much smaller than the QCD backgrounds~\cite{Wardrope:2014kya,deLima:2014dta}
and thus not included in the present analysis.
%

@@

At the generation level, the following loose selection 
cuts are applied to
background events.
%
Each final state particle in the hard process must have $p_T \ge 20$ GeV, and be located
in the central  rapidity
region with
$| \eta | \le 3.0$.
%
In addition, at the matrix-element level
all final state particles must be separated by a minimum $\Delta R_{\mathrm{min}} =0.1$.
%
We have checked that the generator-level cuts are loose enough to have
no influence over the analysis cuts.
%

The LO cross-sections,
number of generated events and the higher-order $K$-factors
applied to the signal and background
samples are collected in Table~\ref{tab:samples}.
%
We see that the $t\bar{t}$ and QCD $4b$ samples are of
the same order of magnitude, however, the former can be efficiently
reduced by using top quark reconstruction criteria.
%
The $bbjj$ cross-section is more than two orders
of magnitude larger as compared to the
$4b$ result, however, as we will show,
due both to combinatorics and to $b$-quark radiation
during the parton shower it can be comparable or larger
than the irreducible QCD background.
%
 For top quark production, only the hadronic final state is generated.
 %

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \small
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Process &  Generator & $N_{\mathrm{evt}}$ & $\sigma_{\mathrm{LO}}$ (pb)  & $K$-factor \\
\hline
\hline
$pp \to hh$ &  {\tt MadGraph5\_aMC@NLO} & 100K & $1.71\times10^{-2}$  &  2.4  (NNLO+NNLL~\cite{deFlorian:2013jea,deFlorian:2015moa}) \\
\hline
\hline
$pp \to b\bar{b}b\bar{b}$ &  {\tt SHERPA}v2.1.1 & 3M &$1.12 \times10^3$  & 1.6 (NLO~\cite{Alwall:2014hca}) \\
$pp \to b\bar{b}jj$ &  {\tt SHERPA}v2.1.1 & 3M & $2.66 \times 10^5$ & 1.3 (NLO~\cite{Alwall:2014hca}) \\
$pp \to jjjj$ &  {\tt SHERPA}v2.1.1 & 3M  & $9.71\times 10^6$ &  0.6 (NLO~\cite{Bern:2011ep})\\
$pp \to t\bar{t}\to b\bar{b}jjjj$ &  {\tt SHERPA}v2.1.1 & 3M & $2.51\times 10^3$   & 1.4 (NNLO+NNLL~\cite{Czakon:2013goa})\\
\hline
\end{tabular}
\caption{\small Details of the signal and background Monte
  Carlo samples generated,
  together with the corresponding generator-level LO cross-sections.
  %
  For top quark production, only the hadronic final state is generated.
  %
We also provide in each case the corresponding inclusive $K$-factor
  that is applied in each case to correctly normalize the distribution to the known
  higher-order results. \label{tab:samples}
} 
\end{center}
\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Modelling of detector resolution}

@@

To account for the finite angular resolution of the calorimeter,
the $\eta$-$\phi$ plane is divided into regions of
$\Delta \eta \times \Delta \phi=0.1$, and each final state particle
which falls in each of these cells is set of have the same $\eta$
and $\phi$ values of the center of the cell.
%
Finally, the particle's energy is recalculated from the smeared $p_T^\prime$,
$\eta^\prime$ and $\phi^\prime$ values to satisfy the mass-shell
condition.

Our modelling of detector simulation has been motivated
to lead to a mass resolution of
reconstructed Higgs candidates of around 10 GeV, consistent
with the hadronic mass resolutions of ATLAS and CMS.
%
In  Fig.~\ref{fig:higgs-mass-resolution} we show the
invariant mass distributions of the leading
Higgs candidate in signal events for the
resolved category (see Sects.~\ref{sec:analysis}
and~\ref{sec:results}), together with
a Gaussian fit,
used to determine the mass resolution.
%
We indeed  find a mass resolution of about 8\%, corresponding
    to around 10 GeV for $m_H=125$ GeV.
%
    For completeness, in Fig.~\ref{fig:higgs-mass-resolution}
    we also show the corresponding
    mass distributions of Higgs candidates
    in the case of $\la n_{\rm PU}\ra=80$ PU events
    subtracted with {\tt SoftKiller} (see Sect.~\ref{sec:pileup}).
    %
    In the case of PU, the mass resolution is degraded down to
    around 11\%, corresponding to 
    14 GeV.
    %

>>>>

We consider the most relevant background
processes that can mimic the
 $hh\to 4b$ signal process.
%
This includes  QCD $4b$ multi-jet production, as well as
QCD $2b2j$ and $4j$ multi-jet samples.
%
The latter can lead to the event being identified
as signal in the case of light or charm
jets being mistagged as $b$-jets.
%
Although the light jet mistag probability is small, we find that
in general the $2b2j$ and $4j$ backgrounds cannot be neglected because
of their large cross-sections and enhancement from combinatorics, that
increase their contribution as compared to a naive estimate.
%
In addition, we find that there is an important contribution from the radiation
of $b$-quarks off light partons during the parton shower, which enhances the contribution
of parton-level $2b2j$ and $4j$ events being identified as signal events.
%

In addition to the QCD multijet processes,
we also generate $t\bar{t}$ samples
in the fully hadronic final state, which lead to a
$2b4j$ signature that can
contribute due to light jet mis-identification.
%
This background processes leads a similar topology that
the corresponding QCD sample.
%
Leptonic decays of top quarks can be removed by requiring
a lepton veto, and thus are not included here.
%
Single Higgs production processes, such as $Z(\to b\bar{b})h(\to b\bar{b})$
and $t\bar{t}h(\to b\bar{b})$, and electroweak backgrounds, such $Z(\to b\bar{b})b\bar{b}$,
are much smaller than the QCD backgrounds~\cite{Wardrope:2014kya,deLima:2014dta}
and thus not included in the present analysis.
%

@@

At the generation level, the following loose selection 
cuts are applied to
background events.
%
Each final state particle in the hard process must have $p_T \ge 20$ GeV, and be located
in the central  rapidity
region with
$| \eta | \le 3.0$.
%
In addition, at the matrix-element level
all final state particles must be separated by a minimum $\Delta R_{\mathrm{min}} =0.1$.
%
We have checked that these generator-level cuts are loose enough to have
no influence over the analysis cuts.
%

The LO cross-sections,
number of generated events and the higher-order $K$-factors
applied to the signal and background
samples are collected in Table~\ref{tab:samples}.
%
We see that the $t\bar{t}$ and QCD $4b$ samples are of
the same order of magnitude, however, the former can be efficiently
reduced by using top quark reconstruction criteria.
%
The $bbjj$ cross-section is more than two orders
of magnitude larger as compared to the
$4b$ result, however, as we will show,
due both to combinatorics and to $b$-quark radiation
during the parton shower it can be comparable or larger
than the irreducible $4b$ QCD background.
%
 For top quark production, only the hadronic final state is generated.
 %

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \small
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Process &  Generator & $N_{\mathrm{evt}}$ & $\sigma_{\mathrm{LO}}$ (pb)  & $K$-factor \\
\hline
\hline
$pp \to hh$ &  {\tt MadGraph5\_aMC@NLO} & 100K & $1.7\cdot10^{-2}$  &  2.4  (NNLO+NNLL~\cite{deFlorian:2013jea,deFlorian:2015moa}) \\
\hline
\hline
$pp \to b\bar{b}b\bar{b}$ &  {\tt SHERPA}v2.1.1 & 3M &$1.1 \cdot10^3$  & 1.6 (NLO~\cite{Alwall:2014hca}) \\
$pp \to b\bar{b}jj$ &  {\tt SHERPA}v2.1.1 & 3M & $2.7 \cdot 10^5$ & 1.3 (NLO~\cite{Alwall:2014hca}) \\
$pp \to jjjj$ &  {\tt SHERPA}v2.1.1 & 3M  & $9.7\cdot 10^6$ &  0.6 (NLO~\cite{Bern:2011ep})\\
$pp \to t\bar{t}\to b\bar{b}jjjj$ &  {\tt SHERPA}v2.1.1 & 3M & $2.5\cdot 10^3$   & 1.4 (NNLO+NNLL~\cite{Czakon:2013goa})\\
\hline
\end{tabular}
\caption{\small Details of the signal and background Monte
  Carlo samples generated,
  together with the corresponding generator-level LO cross-sections.
  %
  For top quark production, only the hadronic final state is generated.
  %
We also provide in each case the corresponding inclusive $K$-factor
  that is applied in each case to correctly normalize the distribution to the known
  higher-order results. \label{tab:samples}
} 
\end{center}
\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Modelling of detector resolution}
\label{sec:detectormodeling}

@@

To account for the finite angular resolution of the calorimeter,
the $\eta$-$\phi$ plane is divided into regions of
$\Delta \eta \times \Delta \phi=0.1$, and each final state particle
which falls in each of these cells is set of have the same $\eta$
and $\phi$ values of the center of the cell.
%
Finally, the particle's energy is recalculated from the smeared $p_T^\prime$,
$\eta^\prime$ and $\phi^\prime$ values to ensure that the resulting
four-momentum is that of a light-like particle, since we neglect all
hadron masses in this analysis.

Our modelling of detector simulation has been motivated
to lead to a mass resolution of
reconstructed Higgs candidates of around 10 GeV, consistent
with the hadronic mass resolutions of ATLAS and CMS~\cite{Aad:2012gxa,Chatrchyan:2013zna,Aad:2014xzb}.
%
In Fig.~\ref{fig:higgs-mass-resolution} we show the
invariant mass distributions of the leading
Higgs candidate in signal events for the
resolved category (see Sects.~\ref{sec:analysis}
and~\ref{sec:results}), together with
a Gaussian fit,
used to determine the mass resolution.
%
We indeed  find a mass resolution of about 8\%, corresponding
    to around 10 GeV for $m_H=125$ GeV.
%
    For completeness, in Fig.~\ref{fig:higgs-mass-resolution}
    we also show the corresponding
    mass distributions of Higgs candidates
    in the case of $\la n_{\rm PU}\ra=80$ PU events
    subtracted with {\tt SoftKiller} (see Sect.~\ref{sec:pileup}).
    %
    In the case of PU, the mass resolution is degraded down to
    around 11\%, corresponding to 
    14 GeV.
    %

### 2015-12-07 paper/sec-results.tex
More work on the paper

First of all, we compare the cross-sections at various
analysis levels for the three categories, for signal and background events,
by providing a detailed cut-flow.
%
We consider all backgrounds enumerated in Sect.~\ref{mcgeneration},
and also discuss how results are modified if only the $4b$
component is kept.
%
At each stage of the cut-flow, we also provide the
the signal significance, $S/\sqrt{B}$, and the signal
over background ratio, $S/B$, corresponding to the
 HL-LHC with
an integrated luminosity of $\mathcal{L}
=3000$ fb$^{-1}$.
%
For this type of analysis,
it is crucial not only to achieve a good signal
significance, $S/\sqrt{B}$, but also a good signal over background ratio $S/B$.
%
If the latter is too small, an unrealistically precise
measurement of the background would be required.
%
In Table~\ref{tab:cutflowdetails}
we describe this cut-flow,
separated into the (exclusive) boosted, intermediate,
    and resolved categories.
    %
    From top to bottom, additional cuts are required in addition
    to the previous ones.
    %
    The various cuts
    are defined as follows:

@@

%
In the boosted category, after all cuts,
one ends
up with more than 1K signal events expected
at the HL-LHC, swamped  however in a almost 1M background events,
leading to $S/\sqrt{B}=1.2$ and $S/B=0.0014$.
%
It should have been possible to somewhat increase this significance
but using more aggressive cuts,
but this is not needed due to the subsequent
optimisation  performed with the MVA.
%
From  Table~\ref{tab:cutflow_noPU_1}
it is also possible to compute the corresponding pre-MVA
expectations for the LHC Run II with
$\mathcal{L}=300$ fb$^{-1}$: one expects
120 signal events, with signal significance dropping down to
$S/\sqrt{B}\sim 0.3$.
%

@@

\subsection{The role of light  and charm jet mistags}

@@

The reason why the  $2b2j$ process
is not negligible as compared to the $4b$ component
is the following.
%
In the boosted category, for example,
the cross-sections at the C1e
level, that is, after the mass-drop tagging but before
$b$-tagging, is a factor 300 larger in the $2b2j$
sample as compared to the $4b$ sample.
%
Now, after $b$-tagging, the naive expectation would
be a suppression of the former by a factor $f_l^2/f_b^2 =3\cdot 10^{-4}$,
since a total of four $b$-tags are required to classify the
event as a Higgs candidate.
%
So the ratio of $2b2j$ over $4b$ should be
around $\simeq 3\%$.
  %
While  we have checked that this expectation is borne
out at parton level, before showering,
we find that  when parton shower effects
are accounted for, due to the radiation of $b\bar{b}$ pairs from the
light jets as well as due to combinatorics, the
number of  $b$ quarks in the  final state is
increased substantially in the $2b2j$ component as compared
to the parton level content,
enhancing the signal efficiency and making
it comparable to the $4b$ component.

>>>>

First of all, we compare the cross-sections at various
stages of the selection
for the three categories, for signal and background events,
by providing a detailed cut-flow.
%
We consider all backgrounds enumerated in Sect.~\ref{mcgeneration},
and also discuss how results are modified if only the $4b$
component is kept.
%
At each stage of the cut-flow, we also provide the
the signal significance, $S/\sqrt{B}$, and the signal
over background ratio, $S/B$, corresponding to the
 HL-LHC with
an integrated luminosity of $\mathcal{L}
=3000$ fb$^{-1}$.
%
For this type of analysis,
it is crucial not only to achieve a good signal
significance, $S/\sqrt{B}$, but also a good signal over background ratio $S/B$.
%
If the latter is too small, an unrealistically precise
measurement of the background would be required.
%
In Table~\ref{tab:cutflowdetails}
we describe this cut-flow,
separated into the (exclusive) boosted, intermediate,
    and resolved categories.
    %
    From top to bottom, additional cuts are required in addition
    to the previous ones.
    %
    The various cuts
    are defined as follows:

@@

%
In the boosted category, after all cuts,
one ends
up with more than 1K signal events expected
at the HL-LHC, swamped  however in a almost 1M background events,
leading to $S/\sqrt{B}=1.2$ and $S/B=0.0014$.
%
The significance in principle could have been increased
by applying tighter selection requirements,
but this is not needed due to the subsequent
optimisation  performed with the MVA.
%
From  Table~\ref{tab:cutflow_noPU_1}
it is also possible to compute the corresponding pre-MVA
expectations for the LHC Run II with
$\mathcal{L}=300$ fb$^{-1}$: one expects
120 signal events, with signal significance dropping down to
$S/\sqrt{B}\sim 0.3$.
%

@@

\subsection{The role of light  and charm jet mis-identification}

@@

The reason why the  $2b2j$ process
is not negligible as compared to the $4b$ component
is the following.
%
In the boosted category, for example,
the cross-sections at the C1e
level, that is, after the mass-drop tagging but before
$b$-tagging, is more than two orders
of magnitude larger in the $2b2j$
sample as compared to the $4b$ sample.
%
Now, after $b$-tagging, the naive expectation would
be a suppression of the former by a factor $f_l^2/f_b^2 =3\cdot 10^{-4}$,
since a total of four $b$-tags are required to classify the
event as a Higgs candidate.
%
So the ratio of $2b2j$ over $4b$ should be
around $\simeq 3\%$.
  %
While  we have checked that this expectation is borne
out at parton level, before showering,
we find that  when parton shower effects
are accounted for, due to the radiation of $b\bar{b}$ pairs from the
light jets as well as due to combinatorics, the
number of  $b$ quarks in the  final state is
increased substantially in the $2b2j$ component as compared
to the parton level content,
enhancing the signal efficiency and making
it comparable to the $4b$ component.
